---
output: rmarkdown::github_document
editor_options: 
  chunk_output_type: console
---
[![Travis-CI Build Status](https://travis-ci.org/hrbrmstr/worldtilegrid.svg?branch=master)](https://travis-ci.org/hrbrmstr/worldtilegrid)

# worldtilegrid [WIP]

A ggplot2 Geom for World Tile Grids

## Description

A "tile grid map" is a cartogram that uses same-sized tiles in approximate, relative positions of each other to represent a world map. The world tile grid relative position reference system used by this 'ggplot2' 'Geom/Stat' was the original work of 'Jon Schwabish' and converted to 'CSV' by 'Maarten Lambrechts'.

- Ref: <https://policyviz.com/2017/10/12/the-world-tile-grid-map/>
- Ref: <http://www.maartenlambrechts.com/2017/10/22/tutorial-a-worldtilegrid-with-ggplot2.html>

## What's Inside The Tin

The following functions are implemented:

```{r install-ex, results='asis', echo = FALSE}
hrbrpkghelpr::install_block()
```

The following _data_ is included/exported:

`wtg`:	World Tile Grid Basemap Data

## Installation

```{r ingredients, results='asis', echo=FALSE, cache=FALSE}
hrbrpkghelpr::describe_ingredients()
```

```{r message=FALSE, warning=FALSE, error=FALSE, include=FALSE}
options(width=120)
```

## Usage

```{r message=FALSE, warning=FALSE, error=FALSE}
library(worldtilegrid)
library(tidyverse)

# current verison
packageVersion("worldtilegrid")

```

### Example (All countries are in the data set)

```{r message=FALSE, warning=FALSE, error=FALSE, fig.width=8, fig.height=8, fig.retina=2}
iso3cs <- worldtilegrid::wtg$alpha.3

set.seed(1)
data_frame(
  ctry = iso3cs,
  `Thing Val` = sample(1000, length(ctry))
) -> xdf

ggplot(xdf, aes(country = ctry, fill = `Thing Val`)) +
  geom_wtg(border_size = 0.5, border_col = "#252a32") +
  geom_text(
    aes(
      label = stat(alpha.2), 
      colour = I(ifelse(`Thing Val` < 500, "white", "black"))
    ), 
    stat = "wtg", size = 2
  ) + 
  coord_equal() +
  viridis::scale_fill_viridis() +
  labs(title = "World Tile Grid") +
  hrbrthemes::theme_ft_rc() +
  theme_enhance_wtg()
```

### Example (Only a few countries are in the data set)

```{r message=FALSE, warning=FALSE, error=FALSE, fig.width=8, fig.height=8, fig.retina=2}
set.seed(1)
data_frame(
  ctry = worldtilegrid::wtg$alpha.3[1:30],
  `Thing Val` = sample(1000, length(ctry))
) -> xdf

ggplot(xdf, aes(country = ctry, fill = `Thing Val`)) +
  geom_wtg() +
  geom_text(
    aes(
      label = stat(alpha.2),
      colour = I(ifelse(`Thing Val` < 400, "#b2b2b2", "#2b2b2b"))
    ), 
    stat="wtg", size=2
  ) + 
  coord_equal() +
  viridis::scale_fill_viridis() +
  labs(title = "World Tile Grid") +
  hrbrthemes::theme_ft_rc() +
  theme_enhance_wtg()
```

### Facet Example (All countries are in the data set)

```{r message=FALSE, warning=FALSE, error=FALSE, fig.width=10, fig.height=6, fig.retina=2}
set.seed(1)
data_frame(
  ctry = worldtilegrid::wtg$alpha.3,
  `Thing Val` = sample(1000, length(ctry)),
  grp = 'Thing One'
) -> xdf1

data_frame(
  ctry = worldtilegrid::wtg$alpha.3,
  `Thing Val` = sample(1000, length(ctry)),
  grp = 'Thing Two'
) -> xdf2

bind_rows(
  xdf1,
  xdf2
) -> xdf

ggplot(xdf, aes(country = ctry, fill = `Thing Val`)) +
  geom_wtg() +
  coord_equal() +
  facet_wrap(~grp) +
  viridis::scale_fill_viridis() +
  labs(title = "World Tile Grid Facets") +
  hrbrthemes::theme_ft_rc() +
  theme_enhance_wtg()
```

### Facet Example (Only a few countries are in the data set)

The geom will fill in the gaps for you:

```{r message=FALSE, warning=FALSE, error=FALSE, fig.width=10, fig.height=6, fig.retina=2}
set.seed(1)
tibble(
  ctry = sample(iso3cs, 40),
  `Thing Val` = sample(1000, length(ctry)),
  grp = 'Thing One'
) -> xdf1

tibble(
  ctry = sample(iso3cs, 40),
  `Thing Val` = sample(1000, length(ctry)),
  grp = 'Thing Two'
) -> xdf2

bind_rows(
  xdf1,
  xdf2
) -> xdf

ggplot(xdf, aes(country = ctry, fill = `Thing Val`)) +
  geom_wtg(border_size = 0.5) +
  coord_equal() +
  facet_wrap(~grp) +
  scale_fill_viridis_c(
    na.value = alpha(hrbrthemes::ft_cols$gray, 2/5), direction = -1,
    option = "magma"
  ) +
  guides(
    fill = guide_colourbar(title.position = "top", title.hjust = 0.5)
  ) +
  labs(title = "World Tile Grid Facets") +
  hrbrthemes::theme_ipsum_rc() +
  theme_enhance_wtg() +
  theme(legend.position = "bottom") +
  theme(legend.key.width = unit(2, "lines"))
```

